\mytitle

### 2016-08-25 pdocs/thesis/article-section-2.tex
fixed vector notations in article section 2
clarifying notations in article section 2

Background
for velocity based formulation shown here is based on  \citet[p.~45-50]{erleben.thesis}.
% pdf page 64
In following section, these formulations will be clarified by simple example using \bullet\ implementation.
Assuming that force $\vec{f}_{true}(t)$, is known the impulse $\vec{J}$
in the time interval $\Delta t $ can we written as
\begin{equation} \label{eq:impulseIntegral}
\vec{J} = \int_{0}^{\Delta t} \vec{f}_{true}(t) dt
\end{equation}
Using Newton's second law of motion $F=ma$ one can solve for the velocity, $\vec{v}^{\Delta t}$
\begin{equation} \label{eq:impulseIntegraWithNewton}
\int_{0}^{\Delta t} m \frac{d\vec{v}}{dt}dt= \int_{0}^{\Delta t} \vec{f}_{true}(t)
\end{equation}
\begin{equation} \label{eq:impulse}
m(\vec{v}^{\, \Delta t} - \vec{v}^{\, 0})=\vec{J}
\end{equation}
where superscripts denote time, i.e. ${v}^{\Delta t}=v(\Delta t)$. Next position can be found
by integrating the velocity.
Derivation of equations of motion is not shown here in detail but results 
for simple contact can be summarized as
equation \ref{eq:eomL} for locations and \ref{eq:eomV} for velocities.
Symbols are summarized in table \ref{tab:eom} and figure \ref{fig:eom}.
\begin{equation} \label{eq:eomL} % pdf page 69
\vec{s}^{\, t+\Delta t} = s^{t}+\Delta t S \vec{u}^{\, t+\Delta t}
\end{equation}

\begin{equation} \label{eq:eomV}
\vec{u}^{\, t+\Delta t} = u^{t}+\Delta t M^{-1}(C N \vec{f}^{\ t+\Delta t} + \vec{f}_{ext}) 
\end{equation}

@@

\begin{equation} \label{eq:eomV2}
\vec{u}^{t+\Delta t} = u^{t}+\Delta t M^{-1}(
 J_{contact}^T \vec{\lambda}_{contact}
+ J_{joint}^T \vec{\lambda}_{joint}
+ \vec{f}_{ext})
\end{equation}

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (O1) at(2,1);
\coordinate (O2) at(2,4);
\coordinate (C) at(2,2);
\draw (0,0) -- (4,0) -- (4,2) -- (0,2) --(0,0);
\draw (2,2) -- (3.5,5) -- (0.5,5) -- (2,2) ;
\node at (3.5,0.4) {$B_1$};
\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
\node at (2.8,4.5) {$B_2$};
\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
\filldraw (C) circle (0.5mm) node[anchor=north west] {$\vec{p}_1$};
\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{11}$};
\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{21}$};
\draw[-latex,thick] (C) -- ++(0,1.4) node[anchor=west] {$\vec{n}_{1}$};
\end{tikzpicture}
\caption{Illustration of nomenclature for equations of motion.}
\label{fig:eom}
\end{figure}

% pdf page 33, notation in typical ODEs
\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l| l|}
\hline
{\bf Symbol} & {\bf Description} \\  \hline
$\vec{r}_i$ & position of center of mass for body i  \\ \hline
$\vec{q}_i$ & orientation as a quaternion for body i  \\ \hline
$\vec{p}_i$ & contact or joint point i  \\ \hline
$\vec{r}_{ki}$ & $\vec{p}_k - \vec{r}_i$  \\ \hline
$\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
$Q_i$ & rotation matrix $Q_i \in  \mathbb{R}^{4 \times 3}$  \\ \hline
$S$ & generalized transformation matrix $ S \in \mathbb{R}^{7n \times 6n}$  \\ \hline
$\vec{v}_i$ & linear velocity of  center of mass for body i  \\ \hline
$\vec{\omega}_i$ & angular velocity of center of mass for body i  \\ \hline
$\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
$M$ & generalized mass matrix $ M \in \mathbb{R}^{6n \times 6n}$  \\ \hline
$C$ & contact condition matrix  $ C \in \mathbb{R}^{6n \times 3K}$ \\ \hline
$N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$\\ \hline
$J_{contact}$ & Jacobian matrix for contacts  \\ \hline
$\lambda_{contact}$ & vector of lagrange multipliers for contacts  \\ \hline
$J_{joint}$ & Jacobian matrix for joints  \\ \hline
$\lambda_{joint}$ & vector of lagrange multipliers for joints  \\ \hline
\end {tabular}
\end{center}
\caption {Nomenclature for equations of motion} \label{tab:eom}
\end {table}

@@

\begin{equation} \label{eq:matrixQi}
Q_i = \frac{1}{2} \left[ \begin{array}{ccc}
-x_i & -y_i & -z_i \\
s_i & z_i & -y_i \\
-z_i & s_i & x_i \\
y_i & -x_i & s_i
\end{array} \right]
\end{equation}

>>>>

Background
for velocity based formulation shown here is based on  \citet[p.~45-50]{erleben.thesis}.
% pdf page 64
In following section, these formulations will be clarified by simple example using \bullet\ implementation.
Assuming that force $\vec{f}_{true}(t)$, is known the impulse $\vec{J}$
in the time interval $\Delta t $ can we written as
\begin{equation} \label{eq:impulseIntegral}
\vec{J} = \int_{0}^{\Delta t} \vec{f}_{true}(t) dt
\end{equation}
Using Newton's second law of motion $F=ma$ one can solve for the velocity, $\vec{v}^{\Delta t}$
\begin{equation} \label{eq:impulseIntegraWithNewton}
\int_{0}^{\Delta t} m \frac{d\vec{v}}{dt}dt= \int_{0}^{\Delta t} \vec{f}_{true}(t)
\end{equation}
\begin{equation} \label{eq:impulse}
m(\vec{v}^{\, \Delta t} - \vec{v}^{\, 0})=\vec{J}
\end{equation}
where superscripts denote time, i.e. ${v}^{\Delta t}=v(\Delta t)$. Next position can be found
by integrating the velocity.
Derivation of equations of motion is not shown here in detail but results 
for simple contact can be summarized as
equation \ref{eq:eomL} for locations and \ref{eq:eomV} for velocities.
Symbols are summarized in table \ref{tab:eom} and figure \ref{fig:eom}.
\begin{equation} \label{eq:eomL} % pdf page 69
\vec{s}^{\, t+\Delta t} = \vec{s}^{\, t}+\Delta t S \vec{u}^{\, t+\Delta t}
\end{equation}

\begin{equation} \label{eq:eomV}
\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(C N \vec{f}^{\ t+\Delta t} + \vec{f}_{ext}) 
\end{equation}

@@

\begin{equation} \label{eq:eomV2}
\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(
 J_{contact}^T \vec{\lambda}_{contact}
+ J_{joint}^T \vec{\lambda}_{joint}
+ \vec{f}_{ext})
\end{equation}

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (O1) at(2,1);
\coordinate (O2) at(2,4);
\coordinate (C) at(2,2);
\draw (0,0) -- (4,0) -- (4,2) -- (0,2) --(0,0);
\draw (2,2) -- (3.5,5) -- (0.5,5) -- (2,2) ;
\node at (3.5,0.4) {$B_1$};
\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
\node at (2.8,4.5) {$B_2$};
\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
\filldraw (C) circle (0.5mm) node[anchor=north west] {$\vec{p}_1$};
\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{11}$};
\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{21}$};
\draw[-latex,thick] (C) -- ++(0,1.4) node[anchor=west] {$\vec{n}_{1}$};
\node[anchor=west] at (4.5,4.5) {
$\vec{r}_i$  =  orientation of body i 
};
\node[anchor=west] at (4.5,4) {$\vec{p}_k $ = orientation of contact point k};
\node[anchor=west] at (4.5,3.5) {$\vec{r}_{ik} $ = $\vec{p}_k - \vec{r}_i $};
\node[anchor=west] at (4.5,3) {$\vec{n}_{k} $ = normal for contact point k };
\end{tikzpicture}
\caption{Illustration of nomenclature for equations of motion for contact.}
\label{fig:eom-contact}
\end{figure}

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (O1) at(1,1);
\coordinate (O2) at(1,3);
\coordinate (C) at(1,2);
\draw (0,0) -- (2,0) -- (2,2) -- (0,2) --(0,0);
\draw (0,2) -- (2,2) -- (2,4) -- (0,4) --(0,2) ;
\node at (0.5,0.4) {$B_1$};
\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
\node at (0.5,3.5) {$B_2$};
\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
\filldraw (C) circle (0.5mm) node[anchor=north west] {$\vec{p}_1$};
\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{11}$};
\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{21}$};
\node[anchor=west] at (4.5,4.5) {
$\vec{r}_i$  =  orientation of body i 
};
\node[anchor=west] at (4.5,4) {$\vec{p}_k $ = orientation of joint point k};
\node[anchor=west] at (4.5,3.5) {$\vec{r}_{ik} $ = $\vec{p}_k - \vec{r}_i $};
\end{tikzpicture}
\caption{Illustration of nomenclature for equations of motion for joint.}
\label{fig:eom-joint}
\end{figure}

% pdf page 33, notation in typical ODEs
\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l| l|}
\hline
{\bf Symbol} & {\bf Description} \\  \hline
$\vec{r}_i$ & position of center of mass for body i  \\ \hline
$\vec{q}_i$ & orientation as a quaternion for body i  (\ref{eq:vectorqi})  \\ \hline
$\vec{p}_i$ & contact or joint point i  \\ \hline
$\vec{r}_{ki}$ & $\vec{p}_k - \vec{r}_i$  \\ \hline
$\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
$Q_i$ & rotation matrix $Q_i \in  \mathbb{R}^{4 \times 3}$ (\ref{eq:matrixQi})  \\ \hline
$S$ & generalized transformation matrix $ S \in \mathbb{R}^{7n \times 6n}$ (\ref{eq:matrixS}) \\ \hline
$\vec{v}_i$ & linear velocity of  center of mass for body i   \\ \hline
$\vec{\omega}_i$ & angular velocity of center of mass for body i  \\ \hline
$\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
$M$ & generalized mass matrix $ M \in \mathbb{R}^{6n \times 6n}$ (\ref{eq:matrixM})\\ \hline
$I_i$ & inertia tensor for body i  \\ \hline
$C$ & contact condition matrix  $ C \in \mathbb{R}^{6n \times 3K}$ \\ \hline
$N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$  (\ref{eq:matrixN}) \\ \hline
$J_{contact}$ & Jacobian matrix for contacts  \\ \hline
$\lambda_{contact}$ & vector of lagrange multipliers for contacts  \\ \hline
$J_{joint}$ & Jacobian matrix for joints  \\ \hline
$\lambda_{joint}$ & vector of lagrange multipliers for joints  \\ \hline
\end {tabular}
\end{center}
\caption {Nomenclature for equations of motion} \label{tab:eom}
\end {table}

@@

\begin{equation} \label{eq:matrixQi} % 4.13 p 50
Q_i = \frac{1}{2} \left[ \begin{array}{ccc}
-x_i & -y_i & -z_i \\
s_i & z_i & -y_i \\
-z_i & s_i & x_i \\
y_i & -x_i & s_i
\end{array} \right]
\end{equation}
